IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-06 paper/appendix.tex
Some changes towards updating the proof.

In the main body of the paper we have only sketched the proof structure
for Lemma~\ref{lemma:rr-tsni-general}.  Here we give it with full
details.

\begin{proof}[Proof of Lemma~\ref{lemma:rr-tsni-general}]
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$.
    By the definition of |erasef il|, we know that |il canFlowTo lcurr|
    where |lcurr| is the label of |it1|.
    In this case, we do not need to take a step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To see that, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified (the scheduling policy only
    modifies the first task). The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1| (can be seen
    by inspecting all reduction rules), they will get filtered
    by |erasef il| in |ic1'|.  Therefore, the |il|-equivalence of the
    task list is guaranteed.
    Lets consider the possible changes to |iS1|:
    Only three reduction interact with |iS1|,
    thus it suffices to consider these cases:
    \begin{description}
      \item[Case \textsc{I-send}]
      A new message triple with label |il'| gets added to the message
      queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
      the triple will get erased.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      In this case, only the queue of
      task |it1| can change, which gets erased.
    \end{description}
    This ensures that $|ic1'|\approx_{|il|}|ic2'|=|ic2|$.
    %    \alphacondition{We need all scheduling policies to not change the order
    %      of any tasks (except for the first one).  Newly spawned task can appear
    %      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.
    Here, there must be a corresponding
    task |it2| in |ic2|,
    such that |it1=it2| (otherwise |ic1| and
    |ic2| could not be |il|-equivalent).
    However, |it2| might not be at the beginning of the task list yet, but
    all tasks occurring before it must get erased by |erasef il|.
    In |ic2|, we can first take some number of steps until |it2| moves
    to the front of the list.
    This is the case regardless of any additional side conditions $P$ on
    rules, because for all of these tasks, it can either take an actual
    step, or it gets dropped by \textsc{I-noStep} (which is always
    possible, as there cannot be an additional side condition on this
    rule).  All tasks that didn't get dropped are still at a label
    that isn't below |il| and thus get erased.

    %    \alphacondition{The scheduling policy must eventually let any task in
    %      the task list evaluate.  In particular, it cannot get stuck when the
    %      first task gets stuck, or keep executing a small number of tasks
    %      exclusively forever (e.g. just execute the first task all the time
    %      if it gets into an infinite loop).}

    Therefore, after |ic2| has potentially executed some number of steps
    to arrive at |ic2''|, we are now in the situation where $|ic1|\approx_{|il|}|ic2''|$, and the first tasks |it1| and |it2|,
    respectively, don't get erased and are thus equivalent.
    The task |it2| can now take exactly the same step as |it1|;  this
    is true even with arbitrary additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted}, since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2|, and thus those predicates evaluate in the same way.
    Thus, we only
    need to argue that the potential differences in |iS1| and |iS2| cannot
    have an influence on the execution (and we know |iS1| and |iS2| are
    |il|-equivalent).
    Again, only sending and receiving will depend on, or change |iS|,
    so we consider all these cases.
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
    \end{description}
  \end{itemize}
\end{proof}

>>>>

In the main body of the paper we have only sketched the proof structure
for Lemma~\ref{lemma:rr-tsni-general}.  Here we give it with full
details, and first state and proof another lemma that will be used in that
proof.

%\begin{lemma}
%  For any task lists |its|, |its'|, and tasks |it|, |it'| we have that
%  if
%  \begin{enumerate}
%    \item |iconf iS (it1, its) .-> iconf iS' (its')|
%    \item $|erase il it1|=|bullet|$
%  \end{enumerate}
%  then
%\end{lemma}

\begin{lemma}
  \label{lemma:high-not-blocking}
  For any task |it|, task lists |its|, store |iS|, and label |il|, if
  $|erase il it|=|bullet|$, then there exists a task list
  |its'| and a store |iS'| such that
  \[
    |iconf iS (it, its)| |.->| |iconf iS' (its, its')|
    \qquad
    |erase il its'|=|nil|
    \qquad
    |erase il iS'|=|erase il iS|
  \]
\end{lemma}
\begin{proof}
  TODO (consideration of all rules that could be applied)
\end{proof}

%\begin{lemma}
%  \label{lemma:erase-context}
%  For any expressions |ie1|, |ie2|, |ie1'|, context |iE1| as well as
%  label |il|, if
%  \[ |ie1= iHole2 iE1 ie1'| \quad \land \quad |erase il ie1 = erase il ie2| \]
%  then there exist |iE2|, |ie2'| such that
%  \[ |ie2= iHole2 iE2 ie2'| \quad \land \quad
%  |erase il ie1' = erase il ie2'|\neq|bullet| \]
%  Similarly, for any
%  expressions |ie1|, |ie2|, |te1'|, context |iE1| as well as
%  label |il|, if
%  \[ |ie1= tHole2 iE1 te1'| \quad \land \quad |erase il ie1 = erase il ie2| \]
%  then there exist |iE2|, |te2'| such that
%  \[ |ie2= tHole2 iE2 te2'| \quad \land \quad
%  |erase il te1' = erase il te2'|\neq|bullet| \]
%\end{lemma}
%\begin{proof}
%  TODO
%\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:rr-tsni-general}]
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.
  Furthermore, let |ic2| be |iconf iS2 its2|.
  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$.
    By the definition of |erasef il|, we know that |il canFlowTo lcurr|
    where |lcurr| is the label of |it1|.
    In this case, we do not need to take a step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To show this, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified (the scheduling policy only
    modifies the first task). The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1| (can be seen
    by inspecting all reduction rules), they will get filtered
    by |erasef il| in |ic1'|.  Therefore, the |il|-equivalence of the
    task list is guaranteed.
    Lets consider the possible changes to |iS1|:
    Only five reduction interact with |iS1|,
    thus it suffices to consider these cases:
    \begin{description}
      \item[Case \textsc{I-send}]
      A new message triple with label |il'| gets added to the message
      queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
      the triple will get erased.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      In this case, only the queue of
      task |it1| can change, which gets erased.
      \item[Case \textsc{I-new}] The newly allocated address has to be at a
      label at least as high as |lcurr|, and will thus be erased.
      \item[Case \textsc{I-write}] Only addresses with a label |il'| above
      |lcurr| can be written, thus the change in |iS1| will get erased.
    \end{description}
    This ensures that $|ic1'|\approx_{|il|}|ic2'|$, as well as
    $|ic2| |.->|^* |ic2'|$ (in zero steps), as claimed.
    %    \alphacondition{We need all scheduling policies to not change the order
    %      of any tasks (except for the first one).  Newly spawned task can appear
    %      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.
    By the definition of |erasef il|, the task list |its2|
    in |ic2| must be of the
    form |its2', it2, its2''| (for some task lists |its2'|, |its2''| and
    some task |it2|)
    where
    \begin{align}
    |erase il its2'| = |nil| \\
    |erase il it2| = |erase il it1| \\
    |erase il its2''| = |erase il its1|
    \end{align}
    (where |nil| is the empty list of tasks).
    Now, intuitively we will first execute a number of steps to process
    the tasks in |its2'| (execute them one step and move them to the back
    of the task list, or drop them if they are done or stuck).  Then, the task
    |it2| can take the same step as |it1|, which will result in a configuration
    |ic2'|
    with the desired properties.
    More formally, we can proceed as follows:

    First, we can apply Lemma~\ref{lemma:high-not-blocking} continuously
    for all the task in |its2'|,
    until we reach a configuration |ic2''=iconf iS2' (it2, its2'', its2''')|
    for some |its2'''| such
    that |erase il its2''' = nil| and |erase il iS2| = |erase il iS2'|.
    We note that |erase il ic1 = erase il ic2''| (by the definition of
    |erasef il|).

    Now, the first task |it2| in |ic2''| is |il|-equivalent to the task |it1|.
    This implies that the two tasks must have the same id, label and
    can only differ in the expression or store if some subexpression
    is of the form |Labeled il' ie|.  In this case, the expression |ie| could
    be different in the two threads if |lcurr canFlowTo il'|.  However, none of the reduction rules
    depend on an expression in that position, and there is never a
    hole in that position
    where evaluation could take place.  Thus, the same rules will syntactically
    match for both task, and we are left to argue that all premises
    evaluate to the same values for |it1| and |it2|, as well as that
    the resulting states |iS1'| and
    |iS2''| are |il|-equivalent.
    The additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted} are not a problem,
    since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2''|, and thus those predicates evaluate in the same way.
    All other premises are either on the threads labels (which are the same),
    or on the state |iS1|, or |iS2'|, respectively.  Because
    |erase il iS1=erase il iS2'|, all of these also evaluate in the same way,
    as can be seen by simply considering all rules that involve or
    change the state:
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
      \item[Case \textsc{I-new}] The newly allocated address can be the same
      for both |it1| and |it2|, thus resulting in |il|-equivalent states.
      \item[Case \textsc{I-write}] By |erase il it1=earse il it2| both tasks
      write the same value, and therefore the resulting states will still
      be |il|-equivalent.
    \end{description}

  \end{itemize}
\end{proof}
